### 2016-06-27 ch05-initvals/initial-values.tex
Small corrections in Ch05 + some requested ones

Parametric identification of \gls{LTI} systems from either input/output data or non-parametric frequency response data, has been well developed as evidenced by published literature~\citep{Pintelon2012,Ljung1999,Schoukens1999,Pintelon1998,Soderstrom1989,Goodwin1977,Brillinger1981,Sanathanan1963,McKelvey2002}.
Many of such parametric identification algorithms, however, require solving non-linear or non-convex optimization problems.
This often involves an iterative algorithm to produce the optimal model.
 The resulting model, however, depends on which initial estimates are used to start up the optimization process: typically one can end up in local minima, leading to models of sub-optimal quality.
 Hence, good initial estimates are essential to obtain a high quality model estimate.
 In this chapter, the \gls{MLE} is used as a proxy for any parametric transfer function estimator, since the \gls{MLE} is commonly used and has desirable properties such as efficiency and consistency.
 Numerous other parametric identification techniques are devoted to the development of parametric plants $G(q^{-1},\theta)$ and parametric noise models  $H(q^{-1},\theta)$, where  $\theta$ is the model parameters vector \citep{Ljung1999,Soderstrom1989,Pintelon2012}.
 Since the initial value generation method are not specifically tailored towards the \gls{MLE}, other parametric identification frameworks are likely to benefit from these improved starting values as well.
 Hence, the presented approach is orthogonal to initialization schemes such as the work of~\citet{vanHerpen2014} that focuses on improving the optimization algorithm, rather than the starting values.

@@

From~\citet[Section 6.3.2.]{Pintelon2012}, in the frequency domain, and for the \glspl{DFT} of the windowed signals, equation \eqref{ABpolys} is of the form
\begin{align}
A(e^{-j\omega_k})\true{Y}(k) = B(e^{-j\omega_k})\true{U}(k) + I(e^{-j\omega_k})
\label{DFTspectra}
\end{align}
with $q^{-1}$ sampled in  $q_k^{-1} = e^{-j\omega_k}$, where $\omega_k = \frac{2\pi k}{N}$ are the \gls{DFT} frequencies, and $I$ is a polynomial of order $\max(N_a,N_b) - 1$, which depends on the difference between the initial and end conditions of the filters $\true{G}$ and $\true{H}$ during the measurement record.

@@

\begin{assumption}
It is assumed that the orders of the polynomials $A$, $B$ and $I$ are known.
\end{assumption}

@@

The maximum likelihood estimate of the parameter vector $\theta$  containing the coefficients of the $A$, $B$ and $I$  polynomials is obtained by solving the optimization problem
\begin{equation}
  \hat{\theta} = \arg\min_\theta V(\theta) \text{.}
\end{equation}
Since the noise $v(t)$ is assumed to be Gaussian, $V(\theta)$ accords to the weighted least squares cost function~\citep[Section 9.11]{Pintelon2012}:
\begin{equation}\label{eq:MLEcf}
V(\theta) = \sum_{k=1}^{N/2-1}\frac{|\varepsilon(k,\theta)|^2}{\sigma_\varepsilon^2(k,\theta)}
\end{equation}
for $N$ measurements and when $\varepsilon$ denotes the error in equation \eqref{DFTspectra},  \emph{viz}.:
\begin{align}
\varepsilon(k,\theta) = A(k,\theta)Y(k) - B(k,\theta)U(k) - I(k,\theta)\text{.}
\end{align}
The variance of the error is of the form:
\begin{equation}\label{eq:sigmaEps}
\begin{split}
\sigma_\varepsilon^2(k,\theta) 
  &=  |A(k,\theta)|^2\sigma_Y^2(k) 
   +  |B(k,\theta)|^2\sigma_U^2(k)
  - 2\RealPart \left( A(k,\theta) \overline{B(k,\theta)} \sigma_{YU}(k) \right)
\end{split}
\end{equation}
where $\sigma_Y^2$ is independent of $k$ for white noise, and the input error variance is assumed to be zero, i.e. $\sigma^2_U=0$, and similarly, the covariance $\sigma_{YU} = 0$.
Here, $\overline{B}$ denotes the complex conjugate of $B$.

@@

\begin{algorithm}
\caption{Levenberg-Marquardt~\citep{Marquardt1963},~\citep[Sec. 9.L.4]{Pintelon2012}}
\label{alg:initvals:levenberg-marquardt}
\begin{algorithmic}[1]
  \Require Cost function \eqref{eq:MLEcf} is rewritten as $V(\theta) = \InnerProductSelf{\epsilon(\theta)}$. 
\Function{LevenbergMarquardt}{$\epsilon(\theta),  \theta^{\mathrm{init}}$}
   \State $\atIter[0]{\theta} \gets \theta^{\mathrm{init}}$
   \Comment{$\atIter[i]{X}$ denotes $X$ at iteration $i$.}
   \For{$i$ \textbf{in} $1 \to \infty$}
      \State $U \, \Sigma \, W^{\TT} \gets \code{svd}\left( \frac{\partial \epsilon(\atIter[i-1]{\theta})}{\partial \theta}\right)$ 
        \Comment{Singular value decomposition of Jacobian}
      % \State $
      %                  \text{Denote }\Sigma \equiv 
      %                  \left[ \begin{matrix} 
      %                  \sigma_1         & \deemph{0} & \deemph{\cdots} &           & \deemph{0}      \\
      %                  \deemph{0}      & \sigma_2    &                   &           &                   \\
      %                  \deemph{\vdots} &              & \ddots            &           & \deemph{\vdots} \\
      %                                    &              &                   & \sigma_n &                   \\
      %                   \deemph{0}      &              & \deemph{\cdots} &           & \deemph{0}      
      %                   \end{matrix}\right]
      %                $
      % \State \textbf{if} $i=1$ \textbf{ then} $\lambda \gets \max(\sigma) / 100$
      \If{$\lambda$ is undefined}
        \State $\lambda \gets \max(\Set{\Sigma_{11}, \ldots, \Sigma_{nn}})/100$
      \EndIf
      \State $\Sigma_{kk} \gets \frac{\Sigma_{kk}}{(\Sigma_{kk}^2 + \lambda^2)} \quad \forall k \in \Set{1, \ldots, n}$
      \State $\delta\theta \gets - W  \Sigma^{\TT}  U^{\TT} \epsilon(\atIter[i-1]{\theta})$
      \State $\atIter[i]{\theta} \gets \frac{(\atIter[i-1]{\theta} + \delta\theta) }{ \norm[2]{\atIter[i-1]{\theta} + \delta\theta} }$
      \If{$V(\atIter[i]{\theta}) > V(\atIter[i-1]{\theta})$} \Comment{Current step is a deterioration}
         \State $\atIter[i]{\theta} \gets \atIter[i-1]{\theta}$ \Comment{Restart loop from previous estimate}
          \State $\lambda \gets 10 \lambda$ 
      \Else
          \State $\lambda \gets 0.4 \lambda$
      \EndIf
      \If {$\norm[2]{\delta\theta} < n_{\theta}  \cdot 10^{-6} \norm[2]{\atIter[i]{\theta}}$}
          \State \textbf{return} $\hat{\theta} \gets \atIter[i]{\theta}$
      \EndIf
   \EndFor
\EndFunction
\end{algorithmic}
\end{algorithm}

@@

\item \emph{Local polynomial method (LPM) with truncation}
\disclaimer{The time-truncated \gls{LPM} has been explained in \secref{sec:nonparametric:truncation} in more detail. The pertinent details are repeated here to aid readability.}
A good estimate of the \gls{FRF} of the chosen system can be obtained via the truncated \gls{LPM}, which is summarized as follows~\citep{Lumori2014TIM} and \secref{sec:nonparametric:truncation} of this thesis.
Here, we repeat the main points briefly.
The \gls{LPM} is first applied to estimate the \gls{FRF} from a full input/output data record of the \gls{SISO} system.
The smooth characteristics of both the exact \gls{FRF} $\true{G}$ and the transient term $T$ allow for application of the \gls{LPM}, leading to a smooth \gls{FRF} estimate with the transient and noise suppressed.

  With reference to a detailed exposition in~\citep{Lumori2014TIM} and \chapref{sec:nonparametric}, the \gls{LPM} utilizes local polynomials to approximate the transfer function and transient contribution (in least squares sense) and, thus, smooth $\true{G}$ and $T$ around a central frequency $\omega_{k}$ in a local frequency band $\omega_{k+r}\in\LocalWindow$.
  In this chapter, we limit the notations to quadratic polynomials in a window with half-width $\numel{W}=3$, such that
\begin{subequations}\label{lpmImplQuadG}
\begin{align}
\true{G}(\omega_{k+r})&\approx \hat G_k + g_{1,k} r + g_{2,k}r^2
\\
T(\omega_{k+r})&\approx T(\omega_k)+t_{1,k}r + t_{2,k}r^2
\label{lpmImplQuadT}
\end{align}
\end{subequations}
where $r \in \LocalShifts{k}$ with $\LocalShifts{k} = \{-\numel{W},-\numel{W}+1,\dots,\numel{W}\}$ and $\numel{W}$ is a tunable parameter.

@@

The \gls{RFIR} estimator is formulated in the time domain. 
It estimates the impulse response of a discrete time system as the minimizer of the following regularized least squares objective function:
\begin{align}\label{eq:hRFIRdef}
\hat g_{\RFIR} &= \argmin_{g} \norm[2]{y - g \conv u }^2 + \sigma^2 g^{\TT}P^{-1}g
\end{align}
where $g$ is the vectorized impulse response $g(t)$, with $t = 0,1,\cdots,n_h - 1$, assuming that the impulse response is $n_g$ samples long. 
Furthermore, $g\conv u$ is the convolution of $g$ with the input signal $u$, and $P\in \mathbb{R}^{n_g\times n_g}$ is the kernel matrix~\citep{Chen2012}. 
Note that \eqref{eq:hRFIRdef} is quadratic in $g$, such that $\hat g_{\RFIR}$ can be computed analytically in a single step.
The kernel matrix $P$ embodies prior knowledge on the system to be estimated. 
Here, the \gls{DCkernel}~\citep{Chen2012,Pillonetto2010} is used. 
Specifically, the element at $t_1,t_2$ of $P$ is given by
\begin{equation}
P(\alpha,\beta)_{t_1,t_2} = e^{-\alpha|t_1 - t_2|}e^{-\beta(t_1 + t_2)/2}
\end{equation}
 for $t_1,t_2 \in \set{0,1,\cdots,n_h - 1}$ which makes the estimated impulse response to be an exponentially decaying function of $t$ with decay rate $\beta^{-1}$, and correlation length $\alpha^{-1}$. 
 The latter \emph{hyperparameters} $\alpha$, $\beta$ and $\sigma$ are determined using Empirical Bayes~\citep{Carlin2000,Gelman2014}, i.e. as the maximizers of the \gls{LML} of the measured output signal~\citep{Chen2012}:
\begin{align}
\mathrm{LML}(y) = -y^{\TT}\Sigma(\alpha,\beta,\sigma)^{-1}y - \log |\Sigma(\alpha,\beta,\sigma)| \text{,}
\end{align}
where $\Sigma(\alpha,\beta, \sigma) = \sigma^2 \I + \phi^{\TT}P(\alpha,\beta)\phi$ and $\phi$ is a Toeplitz matrix constructed with $u$. 
This problem is solved using \mcode{fmincon} in \MATLAB.

@@

\begin{remark}
Note that methods 3) and 4) result in a non-parametric estimate of the transfer function, represented as the \gls{FRF}. 
A corresponding parametric estimate is obtained as follows. 
A parametric estimator (\gls{GTLS}, \gls{BTLS}, and \gls{MLE}) is invoked, where the input spectrum is considered to be $1$ at all frequencies (Dirac in the time domain), and the output spectrum is set equal to the estimated \gls{FRF} $G_{\bullet}(\omega_k)$, with the transient term $I$ set to zero:
\begin{equation}
  U_{\FRF}(\omega_k) \isdef 1 
  \qquad \text{and} \qquad
  Y_{\FRF}(\omega_k)  \isdef G_{\bullet}(\omega_k)
  \text{.}
\end{equation}
It is important to note that invoking the \gls{GTLS}, \gls{BTLS} and \gls{MLE} on the \glspl{FRF} from methods 3) and 4) will yield a different result from applying the \gls{GTLS}, \gls{BTLS} and \gls{MLE} on the raw data. 
This is because the \glspl{FRF} from 3) and 4) have been smoothed and, thus, have a significantly reduced noise variance, but might be biased slightly.
\end{remark}

@@

The success rate $\successRate_{\bullet}$ of an initialization scheme $\bullet$ is expressed mathematically as the probability function
\begin{equation}
\successRate_{\bullet} = 
  \Prob{  \model{\bullet}=\model{\trueSymbol}  }
\text{.}
\end{equation}
In practice, and since the iterative algorithm usually does not reach the local optimum precisely, the above definition is relaxed to the form
\begin{equation}\label{eq:successrateTol}
\successRate_{\bullet} \isdef \Prob{ \norm[2]{\model{\bullet} - \model{\trueSymbol}} < \absoluteTolerance}
\end{equation}
with $\absoluteTolerance$ a numerical tolerance that will be specified later on.

@@

Both systems are low-pass Chebyshev filters which have been generated using the \MATLAB command \mcode[mathescape]{cheby1(2, $\,Q$, 0.05)}.
Their transfer functions are shown in \figref{fig:exampleFRF}.
Their inputs are excited by zero mean white Gaussian noise with unit variance, and the outputs are disturbed by different realizations of white gaussian noise, as in \figref{fig:oesetup}, with a variance $\sigma_v^2$, such that a prescribed \gls{SNR} defined as
\begin{equation}
  \mathrm{SNR} 
    = \frac{\sigma_v}
           {\rms{\true{y}}}
    \approx \frac{\sigma_{v}}
                 {\left\| \true{G} \right\|_2 \rms{\true{u}}}
    =  \frac{\sigma_{v}}
            {\left\| \true{G} \right\|_2}
  \label{eq:SNR-definition}
  \text{.}
\end{equation}
is attained.
$\rms{x} = \sqrt{N^{-1} \sum_t x^2(t)}$ denotes the root-mean-squared value of $x$.

@@

\subsubsection{Simulation}
A Monte Carlo simulation was set up to determine the success rate of the estimator with different initial estimates.
In each run, both the realization of the excitation signal $u$ and the disturbing noise realization ($e$) are varied.
Each run is repeated for the following ranges:
\begin{itemize}
\item of \gls{SNR}: 
              $-20   \unit{dB}$, 
              $-16.7 \unit{dB}$, 
              $-13.3 \unit{dB}$, 
              $-10   \unit{dB}$, 
              $- 7.8 \unit{dB}$, 
              $- 5.6 \unit{dB}$, 
              $- 3.3 \unit{dB}$, 
              $- 1.1 \unit{dB}$, 
              $  1.1 \unit{dB}$, 
              $  3.3 \unit{dB}$, 
              $  5.6 \unit{dB}$, 
              $  7.8 \unit{dB}$, 
              $ 10   \unit{dB}$, 
              $ 13.3 \unit{dB}$, 
              $ 16.7 \unit{dB}$ and
              $ 20   \unit{dB}$
\item of number of samples $N$ is $3\,368\approx 104\timeconst$ or $4\,462\approx 138\timeconst$ for the first system whose quality factor $Q= 6 \unit{dB}$,
\item and of number of samples $N$ is $20\,250\approx 113\timeconst$ or $26\,792\approx 149\timeconst$ for the second system whose quality factor $Q = 20\unit{dB}$.
\end{itemize}
One can, equivalently, express the experiment length as the number $N_{3\unit{dB}}$ of frequency bins that lie within the $3\unit{dB}$ bandwidth of the system.
The $3\unit{dB}$ bandwidth is defined as the frequency band where the magnitude of the transfer function is, at most, $3\unit{dB}$ below its maximum as indicated in \figref{fig:exampleFRF}.
For both systems, $N$ was chosen such that $N_{3\unit{dB}}$ is respectively $36$ or $48$.
For each value-pair $(\SNR, N_{3\mathrm{dB}})$ and both systems, $n_{\mathrm{MC}} = 200$  simulations and estimations were performed.
Each individual simulation yields an estimate of the model based on which one can determine whether the estimation was successful or not.
This allows to empirically estimate the success rate $\successRate$ in equation \eqref{eq:successrateTol} as the proportion of the successes from the $N_{\MC} = 200$ runs.

%\ML{
\subsubsection{Illustrative Example}
In particular, the simulation for each pair $(\mathrm{SNR}, N_{3\mathrm{dB}})$ yields 200 estimates per initial value strategy.
This is illustrated graphically in \figref{fig:single-sim} for the value-pair $(-1.1\unit{dB},48)$ of the system with $Q = 6 \unit{dB}$, where the obtained distances $\norm[2]{\model{\bullet} - \model{\trueSymbol}}$ are shown for the different strategies, and sorted by an ascending norm.
This effectively shows an empirical cumulative distribution function of the distance measure $\norm[2]{\model{\bullet} - \model{\trueSymbol}}$ as sampled by the Monte Carlo algorithm.%}

@@

%\comment{JL: do we have the numerical results to add a remark like: At an SNR of xdB, the respective success rates of $\model{trunc}$ and $\model{existing}$ are $90\%$ and $95\%$. This means that, for that SNR, chances of failure have been reduced by a factor of 2!
%}
%\ML{Yea, shown below!}
%\EG{
  \figref{fig:successRateVS_SNR36N3dB} shows that, typically, for the system with $Q=6\unit{dB}$ when $N_{3\unit{dB}}=48$ and $-1.1 \unit{dB}$ \gls{SNR}, the existing initial values will fail to deliver a good transfer function estimate for one in four trials ($\approx 74\%$ success rate).
However, by using the truncated \gls{LPM}'s initial estimates, the number of failures is reduced fivefold ($\approx 95\%$ success rate) over the existing initial estimates.
Moreover, the \gls{RFIR} initial values lead to poor estimates in less than $2\%$ ($>$ $98\%$ success rate) of the cases, since none were observed in the Monte Carlo simulation.

@@

>>>>

Parametric identification of \gls{LTI} systems from either input/output data or non-parametric frequency response data, has been well developed as evidenced by published literature~\citep{Pintelon2012,Ljung1999,Schoukens1999,Pintelon1998,Soderstrom1989,Goodwin1977,Brillinger1981,Sanathanan1963,McKelvey2002}.
Many of such parametric identification algorithms, however, require solving non-linear or non-convex optimization problems.
This often involves an iterative algorithm to produce the optimal model.
 The resulting model, however, depends on which initial estimates are used to start up the optimization process: typically one can end up in local minima, leading to models of sub-optimal quality.
 Hence, good initial estimates are essential to obtain a high quality model estimate.
 In this chapter, the \gls{MLE} is used as a proxy for any parametric transfer function estimator, since the \gls{MLE} is commonly used and has desirable properties such as efficiency and consistency.
 Numerous other parametric identification techniques are devoted to the development of parametric plants $G(q^{-1},\theta)$ and parametric noise models  $H(q^{-1},\theta)$, where  $\theta$ is the model parameters vector \citep{Ljung1999,Soderstrom1989,Pintelon2012}.
 Since the initial value generation methods are not specifically tailored towards the \gls{MLE}, other parametric identification frameworks are likely to benefit from these improved starting values as well.
 Hence, the presented approach is complementary to initialization schemes such as the work of~\citet{vanHerpen2014} that focuses on improving the optimization algorithm, rather than the starting values.

@@

From~\citet[Section 6.3.2.]{Pintelon2012}, in the frequency domain, and for the \glspl{DFT} of the windowed signals, equation \eqref{ABpolys} is of the form
\begin{align}
A(e^{-j\omega_k})\true{Y}(k) = B(e^{-j\omega_k})\true{U}(k) + T(e^{-j\omega_k})
\label{DFTspectra}
\end{align}
with $q^{-1}$ sampled in  $q_k^{-1} = e^{-j\omega_k}$, where $\omega_k = \frac{2\pi k}{N}$ are the \gls{DFT} frequencies, and $T$ is a polynomial of order $\max(N_a,N_b) - 1$, which depends on the difference between the initial and end conditions of the filters $\true{G}$ and $\true{H}$ during the measurement record.

@@

\begin{assumption}
It is assumed that the orders of the polynomials $A$, $B$ and $T$ are known.
\end{assumption}

@@

The maximum likelihood estimate of the parameter vector $\theta$  containing the coefficients of the $A$, $B$ and $T$  polynomials is obtained by solving the optimization problem
\begin{equation}
  \hat{\theta} = \arg\min_\theta V(\theta) \text{.}
\end{equation}
Since the noise $v(t)$ is assumed to be Gaussian, $V(\theta)$ accords to the weighted least squares cost function~\citep[Section 9.11]{Pintelon2012}:
\begin{equation}\label{eq:MLEcf}
V(\theta) = \sum_{k=1}^{N/2-1}\frac{|\varepsilon(k,\theta)|^2}{\sigma_\varepsilon^2(k,\theta)}
\end{equation}
for $N$ measurements and when $\varepsilon$ denotes the error in equation \eqref{DFTspectra},  \emph{viz}.:
\begin{align}
\varepsilon(k,\theta) = A(k,\theta)Y(k) - B(k,\theta)U(k) - T(k,\theta)\text{.}
\end{align}
The variance of the error is of the form:
\begin{equation}\label{eq:sigmaEps}
\begin{split}
\sigma_\varepsilon^2(k,\theta) 
  &=  |A(k,\theta)|^2\sigma_Y^2(k) 
   +  |B(k,\theta)|^2\sigma_U^2(k)
  - 2\RealPart \left( A(k,\theta) \overline{B(k,\theta)} \sigma_{YU}(k) \right)
\end{split}
\end{equation}
where $\sigma_Y^2$ is independent of $k$ for white noise, and the input error variance is assumed to be zero, i.e. $\sigma^2_U=0$, and similarly, the covariance $\sigma_{YU} = 0$.
Here, $\overline{B}$ denotes the complex conjugate of $B$.

@@

\begin{algorithm}
\caption{Levenberg-Marquardt~\citep{Marquardt1963},~\citep[Sec. 9.L.4]{Pintelon2012}}
\label{alg:initvals:levenberg-marquardt}
\begin{algorithmic}[1]
  \Require Cost function \eqref{eq:MLEcf} is rewritten as $V(\theta) = \InnerProductSelf{\epsilon(\theta)}$. 
\Function{LevenbergMarquardt}{$\epsilon(\theta),  \theta^{\mathrm{init}}$}
   \State $\atIter[0]{\theta} \gets \theta^{\mathrm{init}}$
   \Comment{$\atIter[i]{X}$ denotes $X$ at iteration $i$.}
   \For{$i$ \textbf{in} $1 \to \infty$}
      \State $U \, \Sigma \, W^{\TT} \gets \code{svd}\left( \frac{\partial \epsilon(\atIter[i-1]{\theta})}{\partial \theta}\right)$ 
        \Comment{Singular value decomposition of Jacobian}
      % \State $
      %                  \text{Denote }\Sigma \equiv 
      %                  \left[ \begin{matrix} 
      %                  \sigma_1         & \deemph{0} & \deemph{\cdots} &           & \deemph{0}      \\
      %                  \deemph{0}      & \sigma_2    &                   &           &                   \\
      %                  \deemph{\vdots} &              & \ddots            &           & \deemph{\vdots} \\
      %                                    &              &                   & \sigma_n &                   \\
      %                   \deemph{0}      &              & \deemph{\cdots} &           & \deemph{0}      
      %                   \end{matrix}\right]
      %                $
      % \State \textbf{if} $i=1$ \textbf{ then} $\lambda \gets \max(\sigma) / 100$
      \If{$\lambda$ is undefined}
        \State $\lambda \gets \max(\Set{\Sigma_{11}, \ldots, \Sigma_{\numel{\theta}\numel{\theta}}})/100$
      \EndIf
      \State $\Sigma_{kk} \gets \frac{\Sigma_{kk}}{(\Sigma_{kk}^2 + \lambda^2)} \quad \forall k \in \Set{1, \ldots, \numel{\theta}}$
      \State $\delta\theta \gets - W  \Sigma^{\TT}  U^{\TT} \epsilon(\atIter[i-1]{\theta})$
      \State $\atIter[i]{\theta} \gets \frac{(\atIter[i-1]{\theta} + \delta\theta) }{ \norm[2]{\atIter[i-1]{\theta} + \delta\theta} }$
      \If{$V(\atIter[i]{\theta}) > V(\atIter[i-1]{\theta})$} \Comment{Current step is a deterioration}
         \State $\atIter[i]{\theta} \gets \atIter[i-1]{\theta}$ \Comment{Restart loop from previous estimate}
          \State $\lambda \gets 10 \lambda$ 
      \Else
          \State $\lambda \gets 0.4 \lambda$
      \EndIf
      \If {$\norm[2]{\delta\theta} < n_{\theta}  \cdot 10^{-6} \norm[2]{\atIter[i]{\theta}}$}
          \State \textbf{return} $\hat{\theta} \gets \atIter[i]{\theta}$
      \EndIf
   \EndFor
\EndFunction
\end{algorithmic}
\end{algorithm}

@@

\item \emph{Local polynomial method (LPM) with truncation}
\disclaimer{The time-truncated \gls{LPM} has been explained in \secref{sec:nonparametric:truncation} in more detail. The pertinent details are repeated here to aid readability.}
A good estimate of the \gls{FRF} of the chosen system can be obtained via the truncated \gls{LPM}, which is summarized as follows~\citep{Lumori2014TIM} and \secref{sec:nonparametric:truncation} of this thesis.
The \gls{LPM} is first applied to estimate the \gls{FRF} from a full input/output data record of the \gls{SISO} system.
The smooth characteristics of both the exact \gls{FRF} $\true{G}$ and the transient term $T$ allow for application of the \gls{LPM}, leading to a smooth \gls{FRF} estimate with the transient and noise suppressed.

  With reference to a detailed exposition in~\citep{Lumori2014TIM} and \chapref{sec:nonparametric}, the \gls{LPM} utilizes local polynomials to approximate the transfer function and transient contribution (in least squares sense) and, thus, smooth $\true{G}$ and $T$ around a central frequency $\omega_{k}$ in a local frequency band $\omega_{k+r}\in\LocalWindow$.
  In this chapter, we limit the notations to quadratic polynomials ($\order{B} = \order{T} = 2$) in a window with half-width $\numel{W}=3$, such that
\begin{subequations}\label{lpmImplQuadG}
\begin{align}
\true{G}(\omega_{k+r})&\approx \hat G_k + g_{1,k} r + g_{2,k}r^2
\\
T(\omega_{k+r})&\approx T(\omega_k)+t_{1,k}r + t_{2,k}r^2
\label{lpmImplQuadT}
\end{align}
\end{subequations}
where $r \in \LocalShifts{k}$ with $\LocalShifts{k} = \{-\numel{W},-\numel{W}+1,\dots,\numel{W}\}$.
In general (see \chapref{sec:nonparametric}), $\numel{W}$, $\order{B}$, and $\order{T}$ are tunable parameters.

@@

The \gls{RFIR} estimator is formulated in the time domain. 
It estimates the impulse response of a discrete time system as the minimizer of the following regularized least squares objective function:
\begin{align}\label{eq:hRFIRdef}
\hat g_{\RFIR} &= \argmin_{g} \norm[2]{y - g \conv u }^2 + \sigma^2 g^{\TT}P^{-1}g
\end{align}
where $g$ is the vectorized impulse response $g(t)$, with $t = 0,1,\cdots,n_h - 1$, assuming that the impulse response is $n_g$ samples long. 
Furthermore, $g\conv u$ is the convolution of $g$ with the input signal $u$, and $P\in \mathbb{R}^{n_g\times n_g}$ is the kernel matrix~\citep{Chen2012}. 
Note that \eqref{eq:hRFIRdef} is quadratic in $g$, such that $\hat g_{\RFIR}$ can be computed analytically in a single step.
The kernel matrix $P$ embodies prior knowledge on the system to be estimated. 
Here, the \gls{DCkernel}~\citep{Chen2012,Pillonetto2010} is used. 
Specifically, the element at $t_1,t_2$ of $P$ is given by
\begin{equation}
P(\alpha,\beta)_{t_1,t_2} = e^{-\alpha|t_1 - t_2|}e^{-\beta(t_1 + t_2)/2}
\end{equation}
 for $t_1,t_2 \in \set{0,1,\cdots,n_h - 1}$ which makes the estimated impulse response to be an exponentially decaying function of $t$ with decay rate $\beta^{-1}$, and correlation length $\alpha^{-1}$. 
 The latter \emph{hyperparameters} $\alpha$, $\beta$ and $\sigma$ are determined using Empirical Bayes~\citep{Carlin2000,Gelman2014}, i.e. as the maximizers of the \gls{LML} of the measured output signal~\citep{Chen2012}:
\begin{align}
\mathrm{LML}(y) = -y^{\TT}\Sigma(\alpha,\beta,\sigma)^{-1}y - \log |\Sigma(\alpha,\beta,\sigma)| \text{,}
\end{align}
where $\Sigma(\alpha,\beta, \sigma) = \sigma^2 \I + \phi^{\TT}P(\alpha,\beta)\phi$ and $\phi$ is a Toeplitz matrix constructed with $u$. 
This problem is solved using \mcode{fmincon} in \MATLAB.
%TODO: check convexity of this problem

@@

\begin{remark}
Note that methods 3) and 4) result in a non-parametric estimate of the transfer function, represented as the \gls{FRF}. 
A corresponding parametric estimate is obtained as follows. 
A parametric estimator (\gls{GTLS}, \gls{BTLS}, and \gls{MLE}) is invoked, where the input spectrum is considered to be $1$ at all frequencies (Dirac in the time domain), and the output spectrum is set equal to the estimated \gls{FRF} $G_{\bullet}(\omega_k)$, with the transient term $T$ set to zero:
\begin{equation}
  U_{\FRF}(\omega_k) \isdef 1 
  \qquad \text{and} \qquad
  Y_{\FRF}(\omega_k)  \isdef G_{\bullet}(\omega_k)
  \text{.}
\end{equation}
It is important to note that invoking the \gls{GTLS}, \gls{BTLS} and \gls{MLE} on the \glspl{FRF} from methods 3) and 4) will yield a different result from applying the \gls{GTLS}, \gls{BTLS} and \gls{MLE} on the raw data. 
This is because the \glspl{FRF} from 3) and 4) have been smoothed and, thus, have a significantly reduced noise variance, but might be biased slightly.
\end{remark}

@@

The success rate $\successRate_{\bullet}$ of an initialization scheme $\bullet$ is defined as the probability that it reaches the \emph{best local optimum}, i.e.
\begin{equation}
\successRate_{\bullet} = 
  \Prob{  \model{\bullet} \approx \model{\trueSymbol}  }
\text{.}
\end{equation}
In practice, and since the iterative algorithm usually does not reach the local optimum precisely, the above definition is implemented practically as
\begin{equation}\label{eq:successrateTol}
\successRate_{\bullet} \isdef \Prob{ \norm[2]{\model{\bullet} - \model{\trueSymbol}} < \absoluteTolerance}
\end{equation}
with $\absoluteTolerance$ a numerical tolerance that will be specified later on.

@@

Both systems are low-pass Chebyshev filters which have been generated using the \MATLAB command \mcode[mathescape]{cheby1(2, $\,Q$, 0.05)}.
Their transfer functions are shown in \figref{fig:exampleFRF}.
Their inputs are excited by zero mean white Gaussian noise with unit variance, and the outputs are disturbed by different realizations of white Gaussian noise, as in \figref{fig:oesetup}, with a variance $\sigma_v^2$, such that a prescribed \gls{SNR} defined as
\begin{equation}
  \mathrm{SNR} 
    = \frac{\sigma_v}
           {\rms{\true{y}}}
    \approx \frac{\sigma_{v}}
                 {\left\| \true{G} \right\|_2 \rms{\true{u}}}
    =  \frac{\sigma_{v}}
            {\left\| \true{G} \right\|_2}
  \label{eq:SNR-definition}
  \text{.}
\end{equation}
is attained.
$\rms{x} = \sqrt{N^{-1} \sum_t x^2(t)}$ denotes the root-mean-squared value of $x$.

@@

\subsubsection{Simulation}
A Monte Carlo simulation was set up to determine the success rate of the estimator with different initial estimates.
In each run, both the realization of the excitation signal $u$ and the disturbing noise realization ($e$) are varied.
Each run is repeated for the following ranges:
\begin{itemize}
\item of \gls{SNR}: 
              $-20   \unit{dB}$, 
              $-16.7 \unit{dB}$, 
              $-13.3 \unit{dB}$, 
              $-10   \unit{dB}$, 
              $- 7.8 \unit{dB}$, 
              $- 5.6 \unit{dB}$, 
              $- 3.3 \unit{dB}$, 
              $- 1.1 \unit{dB}$, 
              $  1.1 \unit{dB}$, 
              $  3.3 \unit{dB}$, 
              $  5.6 \unit{dB}$, 
              $  7.8 \unit{dB}$, 
              $ 10   \unit{dB}$, 
              $ 13.3 \unit{dB}$, 
              $ 16.7 \unit{dB}$ and
              $ 20   \unit{dB}$
\item of number of samples $N$ is $3\,368\approx 104\timeconst$ or $4\,462\approx 138\timeconst$ for the first system whose quality factor $Q= 6 \unit{dB}$,
\item and of number of samples $N$ is $20\,250\approx 113\timeconst$ or $26\,792\approx 149\timeconst$ for the second system whose quality factor $Q = 20\unit{dB}$.
\end{itemize}
One can, equivalently, express the experiment length as the number $\nBW$ of frequency bins that lie within the $3\unit{dB}$ bandwidth of the system.
The $3\unit{dB}$ bandwidth is defined as the frequency band where the magnitude of the transfer function is, at most, $3\unit{dB}$ below its maximum as indicated in \figref{fig:exampleFRF}.
For both systems, $N$ was chosen such that $\nBW$ is respectively $36$ or $48$.
For each value-pair $(\SNR, \nBW)$ and both systems, $\nMC = 200$  simulations and estimations were performed.
Each individual simulation yields an estimate of the model based on which one can determine whether the estimation was successful or not.
This allows to empirically estimate the success rate $\successRate$ in equation \eqref{eq:successrateTol} as the proportion of the successes from the $\nMC = 200$ runs.

%\ML{
\subsubsection{Illustrative Example}
In particular, the simulation for each pair $(\mathrm{SNR}, \nBW)$ yields 200 estimates per initial value strategy.
This is illustrated graphically in \figref{fig:single-sim} for the value-pair $(-1.1\unit{dB},48)$ of the system with $Q = 6 \unit{dB}$, where the obtained distances $\norm[2]{\model{\bullet} - \model{\trueSymbol}}$ are shown for the different strategies, and sorted by an ascending norm.
This effectively shows an empirical cumulative distribution function of the distance measure $\norm[2]{\model{\bullet} - \model{\trueSymbol}}$ as sampled by the Monte Carlo algorithm.%}

@@

%\comment{JL: do we have the numerical results to add a remark like: At an SNR of xdB, the respective success rates of $\model{trunc}$ and $\model{existing}$ are $90\%$ and $95\%$. This means that, for that SNR, chances of failure have been reduced by a factor of 2!
%}
%\ML{Yea, shown below!}
%\EG{
  \figref{fig:successRateVS_SNR36N3dB} shows that, typically, for the system with $Q=6\unit{dB}$ when $\nBW=48$ and $-1.1 \unit{dB}$ \gls{SNR}, the existing initial values will fail to deliver a good transfer function estimate for one in four trials ($\approx 74\%$ success rate).
However, by using the truncated \gls{LPM}'s initial estimates, the number of failures is reduced fivefold ($\approx 95\%$ success rate) over the existing initial estimates.
Moreover, the \gls{RFIR} initial values lead to poor estimates in less than $2\%$ ($>$ $98\%$ success rate) of the cases, since none were observed in the Monte Carlo simulation.

@@

%TODO: TM: maybe look into power consumption aspects?
